Symmetric Peer-to-Peer Applications

### 2022-08-21 tpd-22.tex
...

We guide our discussion through the didactic example of Figure~\ref{fig.move},
in which multiple users use the arrow keys to control the direction of a
moving rectangle.
The six instances are connected in a peer-to-peer network, such that some
peers can only communicate indirectly.
%
The application is symmetric in the sense that if any peer presses a key, an
event is broadcast and applied in all peers simultaneously, as if they were a
single mirrored instance.
At any moment, if a peer pauses the application, all peers are guaranteed to
pause exactly at the same frame with the rectangle at the same position.
%
A video

@@

- same keys, same behavior, pause at the same time

we show how a

- sym
- ids
- walk back-forth
- vector clocks
- dynamic net
- re-connect
- late join

>>>>

We guide our discussion through the didactic example of Figure~\ref{fig.move},
in which multiple users use the arrow keys to control the direction of a
moving rectangle.
The six instances are connected in a peer-to-peer network, such that some
peers can only communicate indirectly.
%
The application is symmetric in the sense that if any peer presses a key, a
corresponding event is broadcast and applied in all peers simultaneously, as
if they were a single mirrored instance.
At any moment, if a peer pauses the application, all peers are guaranteed to
pause exactly at the same frame with the rectangle ``bit-identically'' at the
same position.

The main job of the middleware is to deal with the uncertainties of the
network in order to preserve the symmetric behavior across peers.
More specifically, the middleware ensures that all peers
    (i)   meet event deadlines;
    (ii)  advance in sync in real time;
    (iii) can leave and join the network and remain symmetric.
We detail how the middleware preserves these properties further.

\subsection{The Programming API}
\label{sec.tml.api}

The middleware expects to take full control of the application execution in
order to control its timeline, and also communicate events in the network.
For this reason, the basic API is to call an \code{init} function passing a
set of callbacks, which the middleware uses to communicate with the
application, and then call \code{loop} to yield control to the middleware:

{\footnotesize
\begin{verbatim}
int main (void) {
    ...  // initialize the UI
    p2p_init(1, 50, sizeof(G), &G, cb_sim, cb_eff, cb_rec);
    p2p_link("182.168.0.2", 5000, 2);
    p2p_loop();
    p2p_quit();
    ...  // finalize the UI
}
\end{verbatim}
}

The middleware relies on the library \emph{SDL} to deal with the network and
input events, such as timers, mouse clicks and key presses.
The code that follows is a typical:

{\footnotesize
\begin{verbatim}
// preallocated full application state
struct {
    ...
} G;

// executes a simulation step (handles ticks or events)
void cb_sim (p2p_evt evt) {
    switch (evt.id) {
        ...
    }
}

// executes the simulation effects (e.g., redraw the screen)
void cb_eff (int trv) {
    ...
}

// generates events
int cb_rec (SDL_Event* sdl, p2p_evt* evt) {
    switch (sdl->type) {
    	...
    }
}

\end{verbatim}
}

for computer graphics, and
\emph{Arduino} for embedded systems.%
\footnote{SDL: \url{www.libsdl.org}. Arduino: \url{www.arduino.cc}.}
This allows for an easier integration of our middleware with practical systems,
and also reinforces how, in our proposal, programming distributed versions
becomes similar to their local counterparts.

{\footnotesize
\begin{verbatim}
void p2p_init (
    uint8_t me, int port, int fps,
    int mem_n, void* mam_app,
    void (*cb_sim) (p2p_evt),           // simulation callback
    void (*cb_eff) (int trv),           // effects callback
    int (*cb_rec) (SDL_Event*,p2p_evt*) // event recording callback
);
void p2p_link (char* host, int port, uint8_t me);
void p2p_quit (void);
void p2p_loop (void);

void p2p_bcast (uint32_t tick, p2p_evt* evt);
void p2p_dump  (void);

 > freechains '#p2p.md' dislike 3_AE3A1B.. --sign=USR1
 > freechains '#p2p.md' dislike 3_AE3A1B.. --sign=USR2
 > freechains '#p2p.md' dislike 3_AE3A1B.. --sign=USR3
 > freechains-vcs '#p2p.md' checkout p2p.md
 > cat p2p.md
 P2p networking is...
 The [USENET](#usenet.md), ...  <-- no 3rd line
\end{verbatim}
}

%
\begin{itemize}
\item \textbf{Event Deadlines:}
Due to the inherent latency of networks, peers generate events with deliberate
delays so that they reach other peers in time to be applied in sync in a
common deadline.
However, a peer that is locally at time \code{T} may receive an event \code{E}
that should have been applied at time \code{T-t}.
In this case, the middleware automatically rolls back the peer \code{t} units
of time and applies \code{E} at the correct time.
It then fast forwards the peer \code{t} units of time, re-applying all events
that occurred in between \code{T-t} and \code{T}.
%
\item \textbf{Time Synchronization:}
Each peer has its own timeline initiated as soon as the application is
launched.
Also, the internal clock of each peer may diverge over time.
Hence, the timelines will inevitably be out of sync eventually.
The middleware synchronizes clocks continuously as follows:
Every time a

		Each peer also
%
\item[Peer Churn] Arrival and Departure of peers. received events, time sync, all deadlines were missed, departure no problem
\end{itemize}
- time sync

The middleware handles:
- time sync
- disc / conn
- late event

@@

- walk back-forth
- vector clocks
- dynamic net
- re-connect
- late join

We make some assumptions about the network, which are assigned externally to
the middleware:
%
\begin{itemize}
\item Peers have contiguous numeric unique identifiers.
\item Peers identifiers are assigned externally to the middleware.
\item Topology is dynamic, but also assigned externally

\end{itemize}
%
Except for the last item, we claim that the others can be solved externally,
without affecting the middleware.
For instance,
- orthogonal and would not affect the other algorithms in the middleware

, which we claim that could be
solved further:
- peers have unique identifiers
- ids are outside
- topology, dynamic but outside
- fixed maximum nodes
- fixed latency
We claim that Use external protocols
For instance, latency is only a constant, no problem being a variable

- non-malicious nodes
